---
output: pdf_document
---

```{r}

library(readr)
library(ggplot2)

```

```{r}

# read in data
re <- read_csv("./realEstate.csv")

# get rid of useless columns
re <- re[,c(-(1:4),-7,-8,-12,-18,-(27:29),-(31:68))]

# remove observations with missing data
re <- na.omit(re)

# make better var names
names(re) <- gsub(" ","", names(re))

# remove unit labels and dollar signs
re$Price <- parse_number(re$Price)
re$SqrFt <- parse_number(re$SqrFt)

inSqft <- rep(NA,length(re$LotSize))
for (i in 1:length(re$LotSize)) {
  if (grepl("sqft",re$LotSize[i],fixed=TRUE)) {
    inSqft[i] <- TRUE
  } else {
    inSqft[i] <- FALSE
  }
}
re$LotSize <- parse_number(re$LotSize)
for (i in 1:length(re$LotSize)) {
  if (inSqft[i]) {
    re$LotSize[i] <- re$LotSize[i]*(1/43560)
  }
}

re$Bath <- as.factor(re$Bath)
re$Beds <- as.factor(re$Beds)

re$LastSoldFor <- parse_number(re$LastSoldFor)
re$LastTaxAssestment <- parse_number(re$LastTaxAssestment)
re$PriceSqrFt <- parse_number(re$PriceSqrFt)

```

# Visualize Relationships

```{r}

# Square Ft.

ggplot(data=re,aes(x=SqrFt^(1/4),y=log(Price))) +
  geom_point()

# Longitude and Latitude

ggplot(data=re,aes(x=Longitude,y=log(Price))) +
  geom_point()

ggplot(data=re,aes(x=Latitude,y=log(Price))) +
  geom_point()

# Lot Size (acres)

ggplot(data=re,aes(x=log(LotSize),y=log(Price))) +
  geom_point()

# Beds and Baths

ggplot(data=re,aes(x=Beds,y=log(Price))) +
  geom_boxplot()

ggplot(data=re,aes(x=Bath,y=log(Price))) +
  geom_boxplot()

# Year Built

ggplot(data=re,aes(x=YearBuilt,y=log(Price))) +
  geom_point()

# Price Square Ft.

# ggplot(data=re,aes(x=PriceSqrFt,y=Price)) + geom_point()

# Last Sold Year

# ggplot(data=re,aes(x=LastSoldYear,y=log(Price))) + geom_point()

# Last Sold For

ggplot(data=re,aes(x=log(LastSoldFor),y=log(Price))) +
  geom_point()

# Last Tax Assestment

ggplot(data=re,aes(x=log(LastTaxAssestment),y=log(Price))) +
  geom_point()

# Last Tax Year

# ggplot(data=re,aes(x=LastTaxYear,y=log(Price))) + geom_point()

# City

ggplot(data=re,aes(x=City,y=log(Price))) +
  geom_boxplot()

# State

ggplot(data=re,aes(x=State,y=log(Price))) +
  geom_boxplot()

# Zip

# ggplot(data=re,aes(x=Zipcode,y=Price)) + geom_boxplot()

# Days On Trulia

ggplot(data=re,aes(x=DaysOnTrulia,y=log(Price))) +
  geom_point()

```

Consider not using PriceSqrFt because I'm pretty sure it's calculated directly from Price and SqrFt, and we already have that information.

Last Sold Year and Last Tax Year don't have a relationship with log price, which makes quite a bit of sense because house sales and taxes just happen.

Ditto for Days on Trulia.

# Potential Transformations for Linear Models

```{r}

# Remove PriceSqFt, Last Sold Year, Last Tax Year, Days on Trulia

ret <- re[,c(-1,-11,-14,-15,-19)]

# Transformations of total data set

ret$Price <- log(re$Price)
ret$SqrFt <- sqrt(re$SqrFt)
ret$LotSize <- log(re$LotSize)
ret$LastSoldFor <- log(re$LastSoldFor)
ret$LastTaxAssestment <- log(re$LastTaxAssestment)

```

# Potential Subsets of Original Data

```{r}

ca <- re[re$State == "CA",c(-1,-11,-14,-15,-19)]
co <- re[re$State == "CO",c(-1,-11,-14,-15,-19)]
la <- re[re$City == "Los Angeles",c(-1,-11,-14,-15,-19)]
den <- re[re$City == "Denver",c(-1,-11,-14,-15,-19)]

```

# Visualization of Subsets to Determine if Transformations Necessary

```{r}

# California

# SqFt

ggplot(data=ca,aes(x=SqrFt^(1/2),y=log(Price))) +
  geom_point()

# Longitude and Latitude

ggplot(data=ca,aes(x=Longitude,y=log(Price))) +
  geom_point()

ggplot(data=ca,aes(x=Latitude,y=log(Price))) +
  geom_point()

# Lot Size (acres)

ggplot(data=ca,aes(x=log(LotSize),y=log(Price))) +
  geom_point()

# Beds and Baths

ggplot(data=ca,aes(x=Beds,y=log(Price))) +
  geom_boxplot()

ggplot(data=ca,aes(x=Bath,y=log(Price))) +
  geom_boxplot()

# Year Built

ggplot(data=ca,aes(x=YearBuilt,y=log(Price))) +
  geom_point()

# Last Sold For

ggplot(data=ca,aes(x=log(LastSoldFor),y=log(Price))) +
  geom_point()

# Last Tax Assestment

ggplot(data=ca,aes(x=log(LastTaxAssestment),y=log(Price))) +
  geom_point()

# City

ggplot(data=ca,aes(x=City,y=log(Price))) +
  geom_boxplot()

# State

ggplot(data=ca,aes(x=State,y=log(Price))) +
  geom_boxplot()

```

```{r}

# Colorado

# SqFt

ggplot(data=co,aes(x=SqrFt^(1/2),y=log(Price))) +
  geom_point()

# Longitude and Latitude

ggplot(data=co,aes(x=Longitude,y=log(Price))) +
  geom_point()

ggplot(data=co,aes(x=Latitude,y=log(Price))) +
  geom_point()

# Lot Size (acres)

ggplot(data=co,aes(x=log(LotSize),y=log(Price))) +
  geom_point()

# Beds and Baths

ggplot(data=co,aes(x=Beds,y=log(Price))) +
  geom_boxplot()

ggplot(data=co,aes(x=Bath,y=log(Price))) +
  geom_boxplot()

# Year Built

ggplot(data=co,aes(x=YearBuilt,y=log(Price))) +
  geom_point()

# Last Sold For

ggplot(data=co,aes(x=log(LastSoldFor),y=log(Price))) +
  geom_point()

# Last Tax Assestment

ggplot(data=co,aes(x=log(LastTaxAssestment),y=log(Price))) +
  geom_point()

# City

ggplot(data=co,aes(x=City,y=log(Price))) +
  geom_boxplot()

# State

ggplot(data=co,aes(x=State,y=log(Price))) +
  geom_boxplot()

```

```{r}

# Los Angeles



```

```{r}

# Denver



```

# Linear Models via 10-fold CV

Note in the paper that Shrinkage Methods such as Lasso and Ridge will help reduce potential variance increase issues due to multicollinearity. Multicollinearity is probably present here since many variables measurements of the same object (house) at two different times.

```{r}

# California



```

```{r}

# Colorado



```

```{r}

# Los Angeles



```

```{r}

# Denver



```



# Code for Appendix (if it is needed for appendix in slides and paper)

```{r show-code, ref.label = all_labels(), echo = TRUE, eval = FALSE}

```
