---
title: "U.S. Presidential Predictions"
author: "Jacy Werner, Khaled Alketbi, Khalifa Alghaithi"
date: "2024-12-03"
output: powerpoint_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


## Introduction

Throughout U.S. history, presidents have been inaugurated into office from all backgrounds. While this may appear random, we explored historical trends to analyze common components key to a candidate's success. This project focuses on three primary subsets:

- Physical Characteristics
- Economic Situation
- Personal Credibility

Data was collected from historical records (1832-2024) focusing on Democratic and Republican parties.


---

## Motivation

With the 2024 election occurring this semester, we asked: **"Is it possible to predict the outcome of presidential elections and identify factors that contribute to a candidate's success?"**

The project is motivated by a desire to understand the evolving qualifications and characteristics of U.S. presidential candidates.

---

## Methodology

We used logistic regression models to analyze relationships between candidate traits and their likelihood of winning. Data collection challenges included:

- Minimal historical data availability before modern records.
- Influence of opinionated topics like family, wealth, and geography.


---

## Constraints

Some constraints affected the credibility of our data and results:

- Politics is opinionated; voter direction can be influenced by family, wealth, and geography.
- Historical data before modern record-keeping is limited and inconsistent.
- Methods of data storage have evolved, altering the reliability of older data.

---

## Candidates' Height

### 

#### Key Findings:
- **Taller candidates have higher odds of winning.**  
<br>
- **For every 1 cm increase in height, odds increase by about 4.7%.**  
<br>
- **The effect is not statistically significant (p-value = 0.12).**
```{r include=FALSE}
library(dplyr)
library(ggplot2)
library(randomForest)
```

```{r, echo=FALSE}


data <- read.csv("Presidential-Physical-Characteristics (1).csv")

data$Win <- factor(data$Win, levels = c("No", "Yes"))

height_model <- glm(Win ~ Height, data = data, family = "binomial")
data$Predicted_Prob <- predict(height_model, type = "response")

ggplot(data, aes(x = Height, y = Predicted_Prob)) +
  geom_point() +
  labs(title = "Predicted Probability of Winning by Height",
       x = "Height (cm)",
       y = "Predicted Probability") +
  theme_minimal()
```


## Candidates' Age

### 

#### Key Findings:
- **Age has a small positive effect on winning.**  
<br>
- **Odds of winning increase by about 1.5% per year.**  
<br>
- **The effect is Not statistically significant (p-value = 0.538).**

```{r, echo=FALSE}
age_model <- glm(Win ~ Age, data = data, family = "binomial")
data$Predicted_Prob_Age <- predict(age_model, type = "response")

ggplot(data, aes(x = Age, y = Predicted_Prob_Age)) +
  geom_point() +
  labs(title = "Predicted Probability of Winning by Age",
       x = "Age (Years)",
       y = "Predicted Probability") +
  theme_minimal()
```

## Facial Hair

### 

#### Key Findings:
- **Candidates with facial hair have about 18% lower odds of winning.**  
<br>
- **The effect is not statistically significant (p-value = 0.655).**  



```{r, echo=FALSE}

facial_hair_model <- glm(Win ~ Facial.Hair, data = data, family = "binomial")

data$Predicted_Prob_FacialHair <- predict(facial_hair_model, type = "response")

library(ggplot2)
ggplot(data, aes(x = Facial.Hair, y = Predicted_Prob_FacialHair)) +
  geom_boxplot(alpha = 0.7, outlier.shape = 16, outlier.colour = "red") +  
  geom_jitter(width = 0.1, alpha = 0.5, color = "black") + 
  labs(
    title = "Predicted Probability of Winning by Facial Hair",
    x = "Facial Hair (Yes/No)",
    y = "Predicted Probability"
  ) +
  theme_minimal()

```


## Combining All Traits

### 

#### Key Findings:
- **Height has the strongest positive association with winning (4.5% increase per cm).**  
<br>
- **Age and facial hair contribute minimally.**  
<br>
- **None of the traits are statistically significant predictors.**  


```{r, echo=FALSE}
combined_model <- glm(Win ~ Height + Age + Facial.Hair, data = data, family = "binomial")
data$Predicted_Prob_Combined <- predict(combined_model, type = "response")

ggplot(data, aes(x = Age, y = Predicted_Prob_Combined, color = Facial.Hair)) +
  geom_point(aes(size = Height)) +
  labs(title = "Predicted Probability of Winning by Age, Height, and Facial Hair",
       x = "Age (Years)",
       y = "Predicted Probability") +
  theme_minimal() +
  scale_color_manual(values = c("No" = "blue", "Yes" = "red"), name = "Facial Hair")
```


## Comparing Party Preferences


### 

#### Key Findings:
- **For Democrats, Height has a statistically significant positive association with winning. the odds of winning (14.9% per cm).**  
<br>
- **For republicans Physical traits do not significantly influence winning outcomes.**  
  


```{r, echo=FALSE}
data_democrats <- data %>% filter(Party == "Democratic")
data_republicans <- data %>% filter(Party == "Republican")

democrats_model <- glm(Win ~ Height + Age + Facial.Hair, data = data_democrats, family = "binomial")
data_democrats$Predicted_Prob <- predict(democrats_model, type = "response")

republicans_model <- glm(Win ~ Height + Age + Facial.Hair, data = data_republicans, family = "binomial")
data_republicans$Predicted_Prob <- predict(republicans_model, type = "response")

data_combined <- bind_rows(
  data_democrats %>% mutate(Party = "Democratic"),
  data_republicans %>% mutate(Party = "Republican")
)

ggplot(data_combined, aes(x = Age, y = Predicted_Prob, color = Party)) +
  geom_point(aes(size = Height)) +
  labs(title = "Predicted Probability of Winning by Party, Age, and Physical Attributes",
       x = "Age (Years)",
       y = "Predicted Probability") +
  theme_minimal() +
  scale_color_manual(values = c("Democratic" = "blue", "Republican" = "red"))
```


# Credibility

## Candidate Education

### 

#### Key Findings:
- **Candidates with formal education are slightly less likely to win compared to candidates without formal education.**  
<br>
- **The effect is not statistically significant (p = 0.481).**

```{r, echo=FALSE}

education_filtered <- data %>%
  group_by(Year) %>%
  filter(n_distinct(Education) > 1)

education_model <- glm(Win ~ Education, family = "binomial", data = education_filtered)

odds_education <- exp(coef(education_model))

education_filtered$Predicted_Prob_Education <- predict(education_model, type = "response")
ggplot(education_filtered, aes(x = Education, y = Predicted_Prob_Education, color = Education)) +
  geom_jitter(width = 0.1, alpha = 0.6) +
  labs(
    title = "Impact of Education on Winning (Filtered Data)",
    x = "Education",
    y = "Predicted Probability of Winning"
  ) +
  theme_minimal() +
  scale_color_manual(values = c("No" = "red", "Yes" = "blue"))

```


## Candidate Political History

### 

#### Key Findings:
- **The candidates with prior political experience are less likely to win compared to candidates without political experience.**
<br>
- **The effect is not statistically significant (p = 0.451).**


```{r, echo=FALSE}

political_filtered <- data %>%
  group_by(Year) %>%
  filter(n_distinct(Political.History) > 1)

political_model <- glm(Win ~ Political.History, family = "binomial", data = political_filtered)

odds_political <- exp(coef(political_model))


political_filtered$Predicted_Prob_Political <- predict(political_model, type = "response")
ggplot(political_filtered, aes(x = Political.History, y = Predicted_Prob_Political, color = Political.History)) +
  geom_jitter(width = 0.1, alpha = 0.6) +
  labs(
    title = "Impact of Political History on Winning (Filtered Data)",
    x = "Political History",
    y = "Predicted Probability of Winning"
  ) +
  theme_minimal() +
  scale_color_manual(values = c("No" = "red", "Yes" = "blue"))

```

## Candidate Business Ownership

### 

#### Key Findings:
- **The candidates who are business owners are more likely to win compared to non-business owners**
<br>
- **The effect is not statistically significant (p = 0.323).**


```{r, echo=FALSE}

business_filtered <- data %>%
  group_by(Year) %>%
  filter(n_distinct(Business.Owner) > 1)

business_model <- glm(Win ~ Business.Owner, family = "binomial", data = business_filtered)


odds_business <- exp(coef(business_model))

business_filtered$Predicted_Prob_Business <- predict(business_model, type = "response")
ggplot(business_filtered, aes(x = Business.Owner, y = Predicted_Prob_Business, color = Business.Owner)) +
  geom_jitter(width = 0.1, alpha = 0.6) +
  labs(
    title = "Impact of Business Ownership on Winning (Filtered Data)",
    x = "Business Ownership",
    y = "Predicted Probability of Winning"
  ) +
  theme_minimal() +
  scale_color_manual(values = c("No" = "red", "Yes" = "blue"))

```

## Combining All Three Credibility Characteristics

### 

#### Key Findings:

- **The model suggests that candidates with formal education or prior political history are less likely to win.**
<br>
- **None of the three credibility traits—Education, Political History, or Business Ownership—are statistically significant predictors of winning**



```{r, echo=FALSE}

combined_filtered <- data %>%
  group_by(Year) %>%
  filter(n() == 2) %>%
  ungroup()  

categorical_vars <- c("Education", "Political.History", "Business.Owner", "Win")
combined_filtered[categorical_vars] <- lapply(combined_filtered[categorical_vars], factor)

rf_combined_model <- randomForest(
  Win ~ Education + Political.History + Business.Owner,
  data = combined_filtered,
  ntree = 500,
  importance = TRUE
)



importance_df <- data.frame(
  Variable = rownames(importance(rf_combined_model)),
  Importance = importance(rf_combined_model)[, 1]
)



ggplot(importance_df, aes(x = reorder(Variable, Importance), y = Importance)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  coord_flip() +
  labs(
    title = "Variable Importance in Predicting Presidential Win",
    x = "Variable",
    y = "Importance"
  ) +
  theme_minimal()

combined_filtered$Predicted_Prob_RF <- predict(rf_combined_model, newdata = combined_filtered, type = "prob")[, 2]


```

## 

```{r echo=FALSE}
ggplot(combined_filtered, aes(x = Predicted_Prob_RF, fill = Win)) +
  geom_histogram(binwidth = 0.1, alpha = 0.7, position = "identity") +
  labs(
    title = "Predicted Probabilities of Winning (Random Forest Model)",
    x = "Predicted Probability",
    y = "Count",
    fill = "Actual Outcome"
  ) +
  theme_minimal()

```


## Predicting the Party that Will Probably Win in the Next Election

### 

#### Key Findings:
- **The random forest model predicts a winning party error rate of 33.67%.**
<br>
- **The confusion matrix shows that the model classifies Democratic candidates correctly 69.39% of the time and Republican candidates correctly 63.27% of the time **


```{r, echo=FALSE}

data$Party <- as.factor(data$Party)

rf_model <- randomForest(
  Party ~ Height + Facial.Hair + Win + Age + War + GDP.Growth.Rate + Education + Political.History + Business.Owner,
  data = data,
  ntree = 500,
  importance = TRUE
)



importance_df <- data.frame(
  Variable = rownames(importance(rf_model)),
  Importance = importance(rf_model)[, 1]
)


ggplot(importance_df, aes(x = reorder(Variable, Importance), y = Importance)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  coord_flip() +
  labs(
    title = "Variable Importance in Predicting Winning Party",
    x = "Variable",
    y = "Importance"
  ) +
  theme_minimal()

```
