---
title: "U.S. Presidential Predictions"
author: "Jacy Werner, Khaled Alketbi, Khalifa Alghaithi"
date: "2024-12-03"
output: powerpoint_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


## Introduction

Throughout U.S. history, presidents have been inaugurated into office from all backgrounds. While this may appear random, we explored historical trends to analyze common components key to a candidate's success. This project focuses on three primary subsets:

- Physical Characteristics
- Economic Situation
- Personal Credibility

Data was collected from historical records (1832-2024) focusing on Democratic and Republican parties.


---

## Motivation

With the 2024 election occurring this semester, we asked: **"Is it possible to predict the outcome of presidential elections and identify factors that contribute to a candidate's success?"**

The project is motivated by a desire to understand the evolving qualifications and characteristics of U.S. presidential candidates.

---

## Methodology

We used logistic regression models to analyze relationships between candidate traits and their likelihood of winning. Data collection challenges included:

- Minimal historical data availability before modern records.
- Influence of opinionated topics like family, wealth, and geography.


---

## Constraints

Some constraints affected the credibility of our data and results:

- Politics is opinionated; voter direction can be influenced by family, wealth, and geography.
- Historical data before modern record-keeping is limited and inconsistent.
- Methods of data storage have evolved, altering the reliability of older data.

---

## Candidates' Height

### 

#### Key Findings:
- **Taller candidates have higher odds of winning.**  
<br>
- **For every 1 cm increase in height, odds increase by about 4.7%.**  
<br>
- **The effect is not statistically significant (p-value = 0.12).**
```{r include=FALSE}
library(dplyr)
library(ggplot2)
```

```{r, echo=FALSE}


data <- read.csv("Presidential-Physical-Characteristics (1).csv")

data$Win <- factor(data$Win, levels = c("No", "Yes"))

height_model <- glm(Win ~ Height, data = data, family = "binomial")
data$Predicted_Prob <- predict(height_model, type = "response")

ggplot(data, aes(x = Height, y = Predicted_Prob)) +
  geom_point() +
  labs(title = "Predicted Probability of Winning by Height",
       x = "Height (cm)",
       y = "Predicted Probability") +
  theme_minimal()
```


## Candidates' Age

### 

#### Key Findings:
- **Age has a small positive effect on winning.**  
<br>
- **Odds of winning increase by about 1.5% per year.**  
<br>
- **The effect is Not statistically significant (p-value = 0.538).**

```{r, echo=FALSE}
age_model <- glm(Win ~ Age, data = data, family = "binomial")
data$Predicted_Prob_Age <- predict(age_model, type = "response")

ggplot(data, aes(x = Age, y = Predicted_Prob_Age)) +
  geom_point() +
  labs(title = "Predicted Probability of Winning by Age",
       x = "Age (Years)",
       y = "Predicted Probability") +
  theme_minimal()
```

## Facial Hair

### 

#### Key Findings:
- **Candidates with facial hair have about 18% lower odds of winning.**  
<br>
- **The effect is not statistically significant (p-value = 0.655).**  



```{r, echo=FALSE}
facial_hair_model <- glm(Win ~ Facial.Hair, data = data, family = "binomial")


data$Predicted_Prob_FacialHair <- predict(facial_hair_model, type = "response")

ggplot(data, aes(x = Facial.Hair, y = Predicted_Prob_FacialHair)) +
geom_boxplot(alpha = 0.7, outlier.shape = 16, outlier.colour = "red") +  
  geom_jitter(width = 0.1, alpha = 0.5, color = "black") 
  labs(title = "Predicted Probability of Winning by Facial Hair",
       x = "Facial Hair (Yes/No)",
       y = "Predicted Probability") +
  theme_minimal()
```


## Combining All Traits

### 

#### Key Findings:
- **Height has the strongest positive association with winning (4.5% increase per cm).**  
<br>
- **Age and facial hair contribute minimally.**  
<br>
- **None of the traits are statistically significant predictors.**  


```{r, echo=FALSE}
combined_model <- glm(Win ~ Height + Age + Facial.Hair, data = data, family = "binomial")
data$Predicted_Prob_Combined <- predict(combined_model, type = "response")

ggplot(data, aes(x = Age, y = Predicted_Prob_Combined, color = Facial.Hair)) +
  geom_point(aes(size = Height)) +
  labs(title = "Predicted Probability of Winning by Age, Height, and Facial Hair",
       x = "Age (Years)",
       y = "Predicted Probability") +
  theme_minimal() +
  scale_color_manual(values = c("No" = "blue", "Yes" = "red"), name = "Facial Hair")
```


## Comparing Party Preferences


### 

#### Key Findings:
- **For Democrats, Height has a statistically significant positive association with winning. the odds of winning (14.9% per cm).**  
<br>
- **For republicans Physical traits do not significantly influence winning outcomes.**  
  


```{r, echo=FALSE}
data_democrats <- data %>% filter(Party == "Democratic")
data_republicans <- data %>% filter(Party == "Republican")

democrats_model <- glm(Win ~ Height + Age + Facial.Hair, data = data_democrats, family = "binomial")
data_democrats$Predicted_Prob <- predict(democrats_model, type = "response")

republicans_model <- glm(Win ~ Height + Age + Facial.Hair, data = data_republicans, family = "binomial")
data_republicans$Predicted_Prob <- predict(republicans_model, type = "response")

data_combined <- bind_rows(
  data_democrats %>% mutate(Party = "Democratic"),
  data_republicans %>% mutate(Party = "Republican")
)

ggplot(data_combined, aes(x = Age, y = Predicted_Prob, color = Party)) +
  geom_point(aes(size = Height)) +
  labs(title = "Predicted Probability of Winning by Party, Age, and Physical Attributes",
       x = "Age (Years)",
       y = "Predicted Probability") +
  theme_minimal() +
  scale_color_manual(values = c("Democratic" = "blue", "Republican" = "red"))
```