---
output: pdf_document
---

```{r setup, include=FALSE}
library(knitr)
library(tinytex)
# knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(include  = FALSE)
```


```{r}

library(readr)
library(ggplot2)
library(leaps)
library(glmnet)
library(car)
library(pls)
library(dplyr)


library(tidyverse)
library(ISLR)
library(MASS)
library(glmnet)
library(leaps)
library(pls)
library(boot)
library(ggpubr)



```



```{r}

# read in data
re <- read_csv("./realEstate.csv")

# get rid of useless columns
re <- re[,c(-(1:4),-7,-8,-12,-18,-(27:29),-(31:68))]

# remove observations with missing data
re <- na.omit(re)

# make better var names
names(re) <- gsub(" ","", names(re))

# remove unit labels and dollar signs
re$Price <- parse_number(re$Price)
re$SqrFt <- parse_number(re$SqrFt)

inSqft <- rep(NA,length(re$LotSize))
for (i in 1:length(re$LotSize)) {
  if (grepl("sqft",re$LotSize[i],fixed=TRUE)) {
    inSqft[i] <- TRUE
  } else {
    inSqft[i] <- FALSE
  }
}
re$LotSize <- parse_number(re$LotSize)
for (i in 1:length(re$LotSize)) {
  if (inSqft[i]) {
    re$LotSize[i] <- re$LotSize[i]*(1/43560)
  }
}

re$Bath <- as.factor(re$Bath)
re$Beds <- as.factor(re$Beds)

re$LastSoldFor <- parse_number(re$LastSoldFor)
re$LastTaxAssestment <- parse_number(re$LastTaxAssestment)
re$PriceSqrFt <- parse_number(re$PriceSqrFt)

```

# Visualize Relationships

I just did transformations on each original plot instead of making a new plot for each, just in case you're confused why the plots are already using transformed data.

```{r}

# Square Ft.

ggplot(data=re,aes(x=SqrFt^(1/4),y=log(Price))) +
  geom_point()

# Longitude and Latitude

ggplot(data=re,aes(x=Longitude,y=log(Price))) +
  geom_point()

ggplot(data=re,aes(x=Latitude,y=log(Price))) +
  geom_point()

# Lot Size (acres)

ggplot(data=re,aes(x=log(LotSize),y=log(Price))) +
  geom_point()

# Beds and Baths

ggplot(data=re,aes(x=Beds,y=log(Price))) +
  geom_boxplot()

ggplot(data=re,aes(x=Bath,y=log(Price))) +
  geom_boxplot()

# Year Built

ggplot(data=re,aes(x=YearBuilt,y=log(Price))) +
  geom_point()

# Price Square Ft.

ggplot(data=re,aes(x=PriceSqrFt,y=Price)) + geom_point()

# Last Sold Year

ggplot(data=re,aes(x=LastSoldYear,y=log(Price))) + geom_point()

# Last Sold For

ggplot(data=re,aes(x=log(LastSoldFor),y=log(Price))) +
  geom_point()

# Last Tax Assestment

ggplot(data=re,aes(x=log(LastTaxAssestment),y=log(Price))) +
  geom_point()

# Last Tax Year

ggplot(data=re,aes(x=LastTaxYear,y=log(Price))) + geom_point()

# City

ggplot(data=re,aes(x=City,y=log(Price))) +
  geom_boxplot()

# State

ggplot(data=re,aes(x=State,y=log(Price))) +
  geom_boxplot()

# Zip

# ggplot(data=re,aes(x=Zipcode,y=Price)) + geom_boxplot()

# Days On Trulia

ggplot(data=re,aes(x=DaysOnTrulia,y=log(Price))) +
  geom_point()

```

Consider not using PriceSqrFt because I'm pretty sure it's calculated directly from Price and SqrFt, and we already have that information. Actually definitely don't use PriceSqrFt cuz that just undermines everything we're trying to accomplish here.

Last Sold Year and Last Tax Year don't have a relationship with log price, which makes quite a bit of sense because house sales and taxes just happen.

Ditto for Days on Trulia.

# Potential Subsets of Original Data

```{r}

ca <- re[re$State == "CA",c(-1,-10,-15)]
co <- re[re$State == "CO",c(-1,-10,-15)]
la <- re[re$City == "Los Angeles",c(-1,-10,-15)]
den <- re[re$City == "Denver",c(-1,-10,-15)]

```

# Visualization of Subsets to Determine if Transformations Necessary

## California

```{r}

# California

# SqFt

ggplot(data=ca,aes(x=SqrFt^(1/2),y=log(Price))) +
  geom_point()

# Longitude and Latitude

ggplot(data=ca,aes(x=Longitude,y=log(Price))) +
  geom_point()

ggplot(data=ca,aes(x=Latitude,y=log(Price))) +
  geom_point()

# Lot Size (acres)

ggplot(data=ca,aes(x=log(LotSize),y=log(Price))) +
  geom_point()

# Beds and Baths

ggplot(data=ca,aes(x=Beds,y=log(Price))) +
  geom_boxplot()

ggplot(data=ca,aes(x=Bath,y=log(Price))) +
  geom_boxplot()

# Year Built

ggplot(data=ca,aes(x=YearBuilt,y=log(Price))) +
  geom_point()

# Last Sold For

ggplot(data=ca,aes(x=log(LastSoldFor),y=log(Price))) +
  geom_point()

# Last Tax Assestment

ggplot(data=ca,aes(x=log(LastTaxAssestment),y=log(Price))) +
  geom_point()

# City

ggplot(data=ca,aes(x=City,y=log(Price))) +
  geom_boxplot()

# State

ggplot(data=ca,aes(x=State,y=log(Price))) +
  geom_boxplot()

# Other Vars that we didn't consider with all the data, but could in a subset

# Zip

ggplot(data=ca,aes(x=Zipcode,y=Price)) + geom_boxplot()

# Last Tax Year

ggplot(data=ca,aes(x=LastTaxYear,y=log(Price))) + geom_point()

# Last Sold Year

ggplot(data=ca,aes(x=LastSoldYear,y=log(Price))) + geom_point()

# Days on Trulia

ggplot(data=ca,aes(x=DaysOnTrulia,y=log(Price))) +
  geom_point()

```

## Transformations California

```{r}

ca_t <- ca[,c(-7,-14,-15)]
ca_t$Price <- log(ca_t$Price)
ca_t$SqrFt <- sqrt(ca_t$SqrFt)
ca_t$LotSize <- log(ca_t$LotSize)
ca_t$LastSoldFor <- log(ca_t$LastSoldFor)
ca_t$LastTaxAssestment <- log(ca_t$LastTaxAssestment)
ca_t$LastTaxYear <- as.factor(ca_t$LastTaxYear)

```

## Colorado

```{r}

# Colorado

# SqFt

ggplot(data=co,aes(x=SqrFt^(1/2),y=log(Price))) +
  geom_point()

# Longitude and Latitude

ggplot(data=co,aes(x=Longitude,y=log(Price))) +
  geom_point()

ggplot(data=co,aes(x=Latitude,y=log(Price))) +
  geom_point()

# Lot Size (acres)

ggplot(data=co,aes(x=log(LotSize),y=log(Price))) +
  geom_point()

# Beds and Baths

ggplot(data=co,aes(x=Beds,y=log(Price))) +
  geom_boxplot()

ggplot(data=co,aes(x=Bath,y=log(Price))) +
  geom_boxplot()

# Year Built

ggplot(data=co,aes(x=YearBuilt,y=log(Price))) +
  geom_point()

# Last Sold For

ggplot(data=co,aes(x=log(LastSoldFor),y=log(Price))) +
  geom_point()

# Last Tax Assestment

ggplot(data=co,aes(x=log(LastTaxAssestment),y=log(Price))) +
  geom_point()

# City

ggplot(data=co,aes(x=City,y=log(Price))) +
  geom_boxplot()

# State

ggplot(data=co,aes(x=State,y=log(Price))) +
  geom_boxplot()

# Other Vars that we didn't consider with all the data, but could in a subset

# Zip

ggplot(data=co,aes(x=Zipcode,y=Price)) + 
  geom_boxplot()

# Last Tax Year

ggplot(data=co,aes(x=LastTaxYear,y=log(Price))) + 
  geom_point()

# Last Sold Year

ggplot(data=co,aes(x=LastSoldYear,y=log(Price))) + 
  geom_point()

# Days on Trulia

ggplot(data=co,aes(x=DaysOnTrulia,y=log(Price))) +
  geom_point()

```

## Transformations for Colorado

```{r}

co_t <- co[,c(-12,-14,-15)]
co_t$Price <- log(co_t$Price)
co_t$SqrFt <- sqrt(co_t$SqrFt)
co_t$LotSize <- log(co_t$LotSize)
co_t$LastSoldFor <- log(co_t$LastSoldFor)
co_t$LastTaxAssestment <- log(co_t$LastTaxAssestment)

```

## LA

```{r}

# Los Angeles

# SqFt

ggplot(data=la,aes(x=SqrFt^(1/2),y=log(Price))) +
  geom_point()

# Longitude and Latitude

ggplot(data=la,aes(x=Longitude,y=log(Price))) +
  geom_point()

ggplot(data=la,aes(x=Latitude,y=log(Price))) +
  geom_point()

# Lot Size (acres)

ggplot(data=la,aes(x=log(LotSize),y=log(Price))) +
  geom_point()

# Beds and Baths

ggplot(data=la,aes(x=Beds,y=log(Price))) +
  geom_boxplot()

ggplot(data=la,aes(x=Bath,y=log(Price))) +
  geom_boxplot()

# Year Built

ggplot(data=la,aes(x=YearBuilt,y=log(Price))) +
  geom_point()

# Last Sold For

ggplot(data=co,aes(x=log(LastSoldFor),y=log(Price))) +
  geom_point()

# Last Tax Assestment


ggplot(data=la,aes(x=log(LastTaxAssestment),y=log(Price))) +
  geom_point()

# City

ggplot(data=la,aes(x=City,y=log(Price))) +
  geom_boxplot()

# Other Vars that we didn't consider with all the data, but could in a subset

# Zip

ggplot(data=la,aes(x=Zipcode,y=Price)) + geom_boxplot()

# Last Tax Year

ggplot(data=la,aes(x=LastTaxYear,y=log(Price))) + geom_point()

# Last Sold Year

ggplot(data=la,aes(x=LastSoldYear,y=log(Price))) + geom_point()

# Days on Trulia

ggplot(data=la,aes(x=DaysOnTrulia,y=log(Price))) +
  geom_point()

```

## Transformations for LA

```{r}

la_t <- la[,c(-7,-13,-14,-15)]
la_t$Price <- log(la_t$Price)
la_t$SqrFt <- sqrt(la_t$SqrFt)
la_t$LotSize <- log(la_t$LotSize)
la_t$LastSoldFor <- log(la_t$LastSoldFor)
la_t$LastTaxAssestment <- log(la_t$LastTaxAssestment)
la_t$LastTaxYear <- as.factor(la_t$LastTaxYear)

```




## Denver

```{r}

# Denver

# SqFt

dsqr <- ggplot(data=den,aes(x=SqrFt^(1/2),y=log(Price))) +
  geom_point()


# Longitude and Latitude

dlong <- ggplot(data=den,aes(x=Longitude,y=log(Price))) +
  geom_point()

dlat <- ggplot(data=den,aes(x=Latitude,y=log(Price))) +
  geom_point()

# Lot Size (acres)

dlot <- ggplot(data=den,aes(x=log(LotSize),y=log(Price))) +
  geom_point()

# Beds and Baths

dbed <- ggplot(data=den,aes(x=Beds,y=log(Price))) +
  geom_boxplot()

dbat <- ggplot(data=den,aes(x=Bath,y=log(Price))) +
  geom_boxplot()

# Year Built

dyear <- ggplot(data=den,aes(x=YearBuilt,y=log(Price))) +
  geom_point()

# Last Sold For

dlasts <- ggplot(data=den,aes(x=log(LastSoldFor),y=log(Price))) +
  geom_point()

# Last Tax Assestment

dlasttaxa <-  ggplot(data=den,aes(x=log(LastTaxAssestment),y=log(Price))) +
  geom_point()

# City

dcity <- ggplot(data=den,aes(x=City,y=log(Price))) +
  geom_boxplot()

# Other Vars that we didn't consider with all the data, but could in a subset

# Zip

dzip <- ggplot(data=den,aes(x=Zipcode,y=Price)) + geom_boxplot()

# Last Tax Year

dlasttaxy <- ggplot(data=den,aes(x=LastTaxYear,y=log(Price))) + geom_point()

# Last Sold Year

dlastsold <- ggplot(data=den,aes(x=LastSoldYear,y=log(Price))) + geom_point()

# Days on Trulia

dday <- ggplot(data=den,aes(x=DaysOnTrulia,y=log(Price))) +
  geom_point()

ggarrange(dsqr, dlong, dlat, dlot,
           # labels = c("A", "B", "C"),
          ncol = 2, nrow = 2)


ggarrange(dbed, dbat, dyear, dlasts,
           # labels = c("A", "B", "C"),
          ncol = 2, nrow = 2)

ggarrange(dlasttaxa, dcity, dzip, dlasttaxy,
           # labels = c("A", "B", "C"),
          ncol = 2, nrow = 2)


ggarrange(dlastsold, dday, 
           # labels = c("Last Sold year", "Days on Trulia"),
          ncol = 2, nrow = 2)

```

## Transformations for Denver

```{r}

den_t <- den[,c(-12,-13,-14,-15)]
den_t$Price <- log(den_t$Price)
den_t$SqrFt <- sqrt(den_t$SqrFt)
den_t$LotSize <- log(den_t$LotSize)
den_t$LastSoldFor <- log(den_t$LastSoldFor)
den_t$LastTaxAssestment <- log(den_t$LastTaxAssestment)

```

# Linear Models formulated/evaluated via 10-fold CV

Note in the paper that Shrinkage Methods such as Lasso and Ridge will help reduce potential variance increase issues due to multicollinearity. Multicollinearity is probably present here since many variable measurements are of the same house at two different times i.e. Price Now, Price last sold for, last tax adjustment. Although I don't see how multicollinearity would be a problem for prediction exactly, maybe it wouldn't be.

Note for Tomorrow, 12/1/21: Consider train function from caret package. Also consider doing LOOCV seeing as we have lots of time (sorta). Need to figure out how to do best subset selection using LOOCV or 10-fold as model evaluator instead of training RSS. See homework 4 as reference for lasso and ridge, remember that they are parameterized models so need to choose lambda (or whatever its called) by CV, and indeed this is the main analysis. Afterwards use above visualizations to justify addition of polynomial terms to models and test those by LOOCV or 10-fold as well. Afterwards consider step function and spline GAMs if someone hasn't gotten to those already.

Zipcode maybe should be excluded as a predictor because it is perfectly collinear with some other variable (most likely a city). I think this happens because there aren't enough observations in the city that are outside of the zipcode or the zipcode encapsulates the entire city.

I decided (I think) just to do a "best" subset model based on my visualization instead, and a linear regression wtih all the predictors. In the slides we can talk about what a pain in the ass best subset selection is. Frankly, I don't think it's worth it, maybe ever the Lasso and Ridge are just so much easier for basically the same functionality. Or we could do it but idk.

When there is only a few observations for a certain factor level all of them can get left out of the training set, which is problematic because then we can't actually predict on them.



For California I removed Zipcode and Bathroom and this solved the perfect collinearity issues.


### JJ work here!


```{r }
# ca_t #CA state
# la_t #LA city
# den_t # Denver city
# co_t # colorado state


#let start from CA state on Polynomial regression

#Base from the plot above, we would want to do LotSize and year build for our polynomial regression
CA = ca_t

n = nrow(CA)
d <- 10
k <- 10
cv_labels <- sample(seq_len(k), n, replace = T)

cv_mseLS <- rep(NA, d)
cv_mseYB <- rep(NA, d)

for (i in 1:d){
  cv_mseLSi <- rep(NA, k)
  cv_mseYBi <- rep(NA, k)
  
  for (j in 1:k){
    train <- CA[cv_labels != j, ]
    test <- CA[cv_labels == j, ]
    
    fitLS <- lm(Price ~ poly(LotSize, i), data = train)
    fitYB <- lm(Price ~ poly(YearBuilt, i), data = train)
    
    predLS <- predict(fitLS, test)
    cv_mseLSi[j] <- mean((predLS - test$Price)^2)
    
    predYB <- predict(fitYB, test)
    cv_mseYBi[j] <- mean((predYB - test$Price)^2)
    
    
  }
  cv_mseLS[i] <- mean(cv_mseLSi)
  cv_mseYB[i] <- mean(cv_mseYBi)
}

which.min(cv_mseLS)
which.min(cv_mseYB)
# 7 would be the best degree to pick for lot size variable
# 4 would be the best degree to pick for year built variabel

lmLS <- lm(Price ~ poly(LotSize, 7), data = CA)
lmYB <- lm(Price ~ poly(YearBuilt, 4), data = CA)
CA_lmLS_poly <- lmLS
CA_lmYB_poly <- lmYB

fLS <- function(x) {
  polyLS <- cbind(rep(1, dim(poly(x, 7))[1]), poly(x, which.min(cv_mseLS)))
  polyLS %*% lmLS$coefficients
}

fYB <- function(x) {
  polyLS <- cbind(rep(1, dim(poly(x, 4))[1]), poly(x, which.min(cv_mseYB)))
  polyLS %*% lmYB$coefficients
}

ggplot(data=CA,aes(x=LotSize,y=Price)) +
  geom_point() +
  stat_function(fun=fLS,color="red") +
  theme_bw()

ggplot(data=CA,aes(x=YearBuilt,y=Price)) +
  geom_point() +
  stat_function(fun=fYB,color="red") +
  theme_bw()


#start prediction part
erLS <- rep(NA, d)
erYB <- rep(NA, d)
for(i in 2: d){
  CA$LotSize.cut <- cut(CA$LotSize, i)
  fitLS2 <- glm(Price ~ LotSize.cut, data = CA)
  erLS[i] <- cv.glm(CA, fitLS2, K = 10)$delta[1]
  
  CA$YearBuilt.cut <- cut(CA$YearBuilt, i)
  fitYB2 <- glm(Price ~ YearBuilt.cut, data = CA)
  erYB[i] <- cv.glm(CA, fitYB2, K = 10)$delta[1]
}
erLS
erYB

plot(2:10, erLS[-1], type = 'l')
dminLS <- which.min(erLS)
points(dminLS, erLS[dminLS])

plot(2:10, erYB[-1], type = 'l')
dminYB <- which.min(erYB)
points(dminYB, erYB[dminYB])

LS <- CA$LotSize
fitLS3 <- lm(Price ~ cut(LotSize, breaks = 10), data = CA)
rrange = range(LS)
grid <- data.frame(LotSize = seq(rrange[1], rrange[2], length.out = 1000))

ggplot() + 
  geom_point(aes(LotSize,Price),data=CA,alpha=0.2) +
  geom_line(aes(grid$LotSize, predict(fitLS3, grid)), colour = "red") +
  theme_bw()


CA_lmLS_SF <- fitLS3


YB <- CA$YearBuilt
fitYB3 <- lm(Price ~ cut(YearBuilt, breaks = 10), data = CA)
rrange = range(YB)
grid <- data.frame(YearBuilt = seq(rrange[1], rrange[2], length.out = 1000))

ggplot() + 
  geom_point(aes(YearBuilt,Price),data=CA,alpha=0.2) +
  geom_line(aes(grid$YearBuilt, predict(fitYB3, grid)), colour = "red") +
  theme_bw()

CA_lmYB_SF <- fitYB3



```



```{r }
#Now we got CA, let do Colroado or CO
CO <- co_t
n = nrow(CO)
d <- 10
k <- 10
cv_labels <- sample(seq_len(k), n, replace = T)

cv_mseLS <- rep(NA, d)
cv_mseYB <- rep(NA, d)

for (i in 1:d){
  cv_mseLSi <- rep(NA, k)
  cv_mseYBi <- rep(NA, k)
  
  for (j in 1:k){
    train <- CO[cv_labels != j, ]
    test <- CO[cv_labels == j, ]
    
    fitLS <- lm(Price ~ poly(LotSize, i), data = train)
    fitYB <- lm(Price ~ poly(YearBuilt, i), data = train)
    
    predLS <- predict(fitLS, test)
    cv_mseLSi[j] <- mean((predLS - test$Price)^2)
    
    predYB <- predict(fitYB, test)
    cv_mseYBi[j] <- mean((predYB - test$Price)^2)
    
    
  }
  cv_mseLS[i] <- mean(cv_mseLSi)
  cv_mseYB[i] <- mean(cv_mseYBi)
}

which.min(cv_mseLS)
which.min(cv_mseYB)
# 2 would be the best degree to pick for lot size variable
# 7 would be the best degree to pick for year built variabel

lmLS <- lm(Price ~ poly(LotSize, 2), data = CO)
lmYB <- lm(Price ~ poly(YearBuilt, 7), data = CO)

CO_lmLS_poly <- lmLS
CO_lmYB_poly <- lmYB

fLS <- function(x) {
  polyLS <- cbind(rep(1, dim(poly(x, 2))[1]), poly(x, 2))
  polyLS %*% lmLS$coefficients
}

fYB <- function(x) {
  polyLS <- cbind(rep(1, dim(poly(x, 7))[1]), poly(x, 7))
  polyLS %*% lmYB$coefficients
}

ggplot(data=CO,aes(x=LotSize,y=Price)) +
  geom_point() +
  stat_function(fun=fLS,color="red") +
  theme_bw()

ggplot(data=CO,aes(x=YearBuilt,y=Price)) +
  geom_point() +
  stat_function(fun=fYB,color="red") +
  theme_bw()


#start prediction part
erLS <- rep(NA, d)
erYB <- rep(NA, d)
for(i in 2: d){
  CO$LotSize.cut <- cut(CO$LotSize, i)
  fitLS2 <- glm(Price ~ LotSize.cut, data = CO)
  erLS[i] <- cv.glm(CO, fitLS2, K = 10)$delta[1]
  
  CO$YearBuilt.cut <- cut(CO$YearBuilt, i)
  fitYB2 <- glm(Price ~ YearBuilt.cut, data = CO)
  erYB[i] <- cv.glm(CO, fitYB2, K = 10)$delta[1]
}
erLS
erYB

plot(2:10, erLS[-1], type = 'l')
dminLS <- which.min(erLS)
points(dminLS, erLS[dminLS])

plot(2:10, erYB[-1], type = 'l')
dminYB <- which.min(erYB)
points(dminYB, erYB[dminYB])

LS <- CO$LotSize
fitLS3 <- lm(Price ~ cut(LotSize, breaks = 10), data = CO)
rrange = range(LS)
grid <- data.frame(LotSize = seq(rrange[1], rrange[2], length.out = 1000))

ggplot() + 
  geom_point(aes(LotSize,Price),data=CO,alpha=0.2) +
  geom_line(aes(grid$LotSize, predict(fitLS3, grid)), colour = "red") +
  theme_bw()



YB <- CO$YearBuilt
fitLS3 <- lm(Price ~ cut(YearBuilt, breaks = 10), data = CO)
rrange = range(YB)
grid <- data.frame(YearBuilt = seq(rrange[1], rrange[2], length.out = 1000))

ggplot() + 
  geom_point(aes(YearBuilt,Price),data=CO,alpha=0.2) +
  geom_line(aes(grid$YearBuilt, predict(fitLS3, grid)), colour = "red") +
  theme_bw()




```

```{r }
# Sub it down to small city
Denver <- den_t

n = nrow(Denver)
d <- 10
k <- 10
cv_labels <- sample(seq_len(k), n, replace = T)

cv_mseLS <- rep(NA, d)
cv_mseYB <- rep(NA, d)

for (i in 1:d){
  cv_mseLSi <- rep(NA, k)
  cv_mseYBi <- rep(NA, k)
  
  for (j in 1:k){
    train <- Denver[cv_labels != j, ]
    test <- Denver[cv_labels == j, ]
    
    fitLS <- lm(Price ~ poly(LotSize, i), data = train)
    fitYB <- lm(Price ~ poly(YearBuilt, i), data = train)
    
    predLS <- predict(fitLS, test)
    cv_mseLSi[j] <- mean((predLS - test$Price)^2)
    
    predYB <- predict(fitYB, test)
    cv_mseYBi[j] <- mean((predYB - test$Price)^2)
    
    
  }
  cv_mseLS[i] <- mean(cv_mseLSi)
  cv_mseYB[i] <- mean(cv_mseYBi)
}

which.min(cv_mseLS)
which.min(cv_mseYB)


#degree for both year built and lot size are 7

lmLS <- lm(Price ~ poly(LotSize, 7), data = Denver)
lmYB <- lm(Price ~ poly(YearBuilt, 7), data = Denver)

DN_lmLS_poly <- lmLS
DN_lmYB_poly <- lmYB


fLS <- function(x) {
  polyLS <- cbind(rep(1, dim(poly(x, 7))[1]), poly(x, 7))
  polyLS %*% lmLS$coefficients
}

fYB <- function(x) {
  polyLS <- cbind(rep(1, dim(poly(x, 7))[1]), poly(x, 7))
  polyLS %*% lmYB$coefficients
}

ggplot(data=Denver,aes(x=LotSize,y=Price)) +
  geom_point() +
  stat_function(fun=fLS,color="red") +
  theme_bw()

ggplot(data=Denver,aes(x=YearBuilt,y=Price)) +
  geom_point() +
  stat_function(fun=fYB,color="red") +
  theme_bw()


#start prediction part
erLS <- rep(NA, d)
erYB <- rep(NA, d)
for(i in 2: d){
  Denver$LotSize.cut <- cut(Denver$LotSize, i)
  fitLS2 <- glm(Price ~ LotSize.cut, data = Denver)
  erLS[i] <- cv.glm(Denver, fitLS2, K = 10)$delta[1]
  
  Denver$YearBuilt.cut <- cut(Denver$YearBuilt, i)
  fitYB2 <- glm(Price ~ YearBuilt.cut, data = Denver)
  erYB[i] <- cv.glm(Denver, fitYB2, K = 10)$delta[1]
}
erLS
erYB

plot(2:10, erLS[-1], type = 'l')
dminLS <- which.min(erLS)
points(dminLS, erLS[dminLS])

plot(2:10, erYB[-1], type = 'l')
dminYB <- which.min(erYB)
points(dminYB, erYB[dminYB])


LS <- Denver$LotSize
fitLS3 <- lm(Price ~ cut(LotSize, breaks = 10), data = Denver)
rrange = range(LS)
grid <- data.frame(LotSize = seq(rrange[1], rrange[2], length.out = 1000))

ggplot() + 
  geom_point(aes(LotSize,Price),data=Denver,alpha=0.2) +
  geom_line(aes(grid$LotSize, predict(fitLS3, grid)), colour = "red") +
  theme_bw()



YB <- Denver$YearBuilt
fitLS3 <- lm(Price ~ cut(YearBuilt, breaks = 4), data = Denver)
rrange = range(YB)
grid <- data.frame(YearBuilt = seq(rrange[1], rrange[2], length.out = 1000))

ggplot() + 
  geom_point(aes(YearBuilt,Price),data=Denver,alpha=0.2) +
  geom_line(aes(grid$YearBuilt, predict(fitLS3, grid)), colour = "red") +
  theme_bw()


```

```{r }
# City LA
LA <- la_t

n = nrow(LA)
d <- 10
k <- 10
cv_labels <- sample(seq_len(k), n, replace = T)

cv_mseLS <- rep(NA, d)
cv_mseYB <- rep(NA, d)

for (i in 1:d){
  cv_mseLSi <- rep(NA, k)
  cv_mseYBi <- rep(NA, k)
  
  for (j in 1:k){
    train <- LA[cv_labels != j, ]
    test <- LA[cv_labels == j, ]
    
    fitLS <- lm(Price ~ poly(LotSize, i), data = train)
    fitYB <- lm(Price ~ poly(YearBuilt, i), data = train)
    
    predLS <- predict(fitLS, test)
    cv_mseLSi[j] <- mean((predLS - test$Price)^2)
    
    predYB <- predict(fitYB, test)
    cv_mseYBi[j] <- mean((predYB - test$Price)^2)
    
    
  }
  cv_mseLS[i] <- mean(cv_mseLSi)
  cv_mseYB[i] <- mean(cv_mseYBi)
}

which.min(cv_mseLS)
which.min(cv_mseYB)


#degree for lot size is 6 and year built is 5

lmLS <- lm(Price ~ poly(LotSize, 6), data = LA)
lmYB <- lm(Price ~ poly(YearBuilt, 5), data = LA)


LA_lmLS_poly <- lmLS
LA_lmYB_poly <- lmYB

fLS <- function(x) {
  polyLS <- cbind(rep(1, dim(poly(x, 6))[1]), poly(x, 6))
  polyLS %*% lmLS$coefficients
}

fYB <- function(x) {
  polyLS <- cbind(rep(1, dim(poly(x, 5))[1]), poly(x, 5))
  polyLS %*% lmYB$coefficients
}

ggplot(data=LA,aes(x=LotSize,y=Price)) +
  geom_point() +
  stat_function(fun=fLS,color="red") +
  theme_bw()

ggplot(data=LA,aes(x=YearBuilt,y=Price)) +
  geom_point() +
  stat_function(fun=fYB,color="red") +
  theme_bw()



#start prediction part
erLS <- rep(NA, d)
# erYB <- rep(NA, d)
for(i in 2: d){
  LA$LotSize.cut <- cut(LA$LotSize, i)
  fitLS2 <- glm(Price ~ LotSize.cut, data = LA)
  erLS[i] <- cv.glm(LA, fitLS2, K = 10)$delta[1]
  # 
  # LA$YearBuilt.cut <- cut(LA$YearBuilt, i)
  # fitYB2 <- glm(Price ~ YearBuilt.cut, data = LA)
  # erYB[i] <- cv.glm(LA, fitYB2, K = 10)$delta[1]
}
erLS
# erYB




plot(2:10, erLS[-1], type = 'l')
dminLS <- which.min(erLS)
points(dminLS, erLS[dminLS])

# plot(2:10, erYB[-1], type = 'l')
# dminYB <- which.min(erYB)
# points(dminYB, erYB[dminYB])


LS <- LA$LotSize
fitLS3 <- lm(Price ~ cut(LotSize, breaks = 3), data = LA)
rrange = range(LS)
grid <- data.frame(LotSize = seq(rrange[1], rrange[2], length.out = 1000))

ggplot() + 
  geom_point(aes(LotSize,Price),data=LA,alpha=0.2) +
  geom_line(aes(grid$LotSize, predict(fitLS3, grid)), colour = "red") +
  theme_bw()



YB <- LA$YearBuilt
fitLS3 <- lm(Price ~ cut(YearBuilt, breaks = 4), data = LA)
rrange = range(YB)
grid <- data.frame(YearBuilt = seq(rrange[1], rrange[2], length.out = 1000))

ggplot() +
  geom_point(aes(YearBuilt,Price),data=LA,alpha=0.2) +
  geom_line(aes(grid$YearBuilt, predict(fitLS3, grid)), colour = "red") +
  theme_bw()



```


```{r }
#prediction
#starting from CA, CO then denver and LA

set.seed(420)
n <- nrow(CA)
catest <- seq_len(n) %in% sample(seq_len(n),round(0.7*n))
n <- nrow(CO)
cotest <- seq_len(n) %in% sample(seq_len(n),round(0.7*n))
n <- nrow(Denver)
dntest <- seq_len(n) %in% sample(seq_len(n),round(0.7*n))
n <- nrow(LA)
latest <- seq_len(n) %in% sample(seq_len(n),round(0.7*n))

pred_polyCA_LS <- predict(CA_lmLS_poly, CA[catest, 5])
pred_polyCA_YB <- predict(CA_lmYB_poly, CA[catest, 7])
# pred_sfCA_LS <- predict(CA_lmLS_SF, CA[catest, 5])
# pred_sfCA_YB <- predict(CA_lmYB_SF, CA[catest, 7])

pred_polyCO_LS <- predict(CO_lmLS_poly, CO[cotest, 5])
pred_polyCO_YB <- predict(CO_lmYB_poly, CO[cotest, 8])

pred_polyDN_LS <- predict(DN_lmLS_poly, Denver[dntest, 5])
pred_polyDN_YB <- predict(DN_lmYB_poly, Denver[dntest, 8])

pred_polyLA_LS <- predict(LA_lmLS_poly, LA[latest, 5])
pred_polyLA_YB <- predict(LA_lmYB_poly, LA[latest, 7])

#getting MSE

MSE_CALS <- mean((CA[catest,]$Price - pred_polyCA_LS)^2)
MSE_CAYB <- mean((CA[catest,]$Price - pred_polyCA_YB)^2)

MSE_COLS <- mean((CO[cotest,]$Price - pred_polyCO_LS)^2)
MSE_COYB <- mean((CO[cotest,]$Price - pred_polyCO_YB)^2)

MSE_DNLS <- mean((Denver[dntest,]$Price - pred_polyDN_LS)^2)
MSE_DNYB <- mean((Denver[dntest,]$Price - pred_polyDN_YB)^2)

MSE_LALS <- mean((LA[latest,]$Price - pred_polyLA_LS)^2)
MSE_LAYB <- mean((LA[latest,]$Price - pred_polyLA_YB)^2)



```



### JJ work end!!!






















<!-- ```{r} -->

<!-- set.seed(420) -->

<!-- # California -->

<!-- n <- nrow(ca) -->
<!-- trn <- seq_len(n) %in% sample(seq_len(n),round(0.7*n)) -->

<!-- n <- length(trn[trn==TRUE]) -->
<!-- folds <- sample(seq_len(k), n, replace = k != n) -->

<!-- # Base Model -->

<!-- base_lm_ca <- lm(Price~.,data=ca_t[trn,c(-6,-12)]) -->
<!-- aliasMatrixCA <- alias(base_lm_ca)$Complete -->

<!-- pred_base_ca <- predict(base_lm_ca,ca_t[!trn,c(-6,-12)]) -->
<!-- mse_base_ca <- mean((ca_t[!trn,c(-6,-12)]$Price - pred_base_ca)^2) -->
<!-- mse_base_ca -->

<!-- # Ridge -->

<!-- cv_ridge_ca <- cv.glmnet(model.matrix(Price~.,data=ca_t)[trn,-1],unlist(ca_t[trn,1]),alpha=0,lambda=lambda,foldid=folds) -->

<!-- ridge_trn_ca <- glmnet(model.matrix(Price~.,data=ca_t)[trn,-1],unlist(ca_t[trn,1]),alpha=0,lambda=cv_ridge_ca$lambda.min) -->
<!-- pred_ridge_ca <- predict(ridge_trn_ca, model.matrix(Price~.,data=ca_t)[!trn,-1]) -->

<!-- mse_ridge_ca <- mean((ca_t[!trn,]$Price - pred_ridge_ca)^2) -->
<!-- mse_ridge_ca -->

<!-- # Lasso -->

<!-- cv_lasso_ca <- cv.glmnet(model.matrix(Price~.,data=ca_t)[trn,-1],unlist(ca_t[trn,1]),alpha=1,lambda=lambda,foldid=folds) -->

<!-- lasso_trn_ca <- glmnet(model.matrix(Price~.,data=ca_t)[trn,-1],unlist(ca_t[trn,1]),alpha=1,lambda=cv_lasso_ca$lambda.min) -->
<!-- pred_lasso_ca <- predict(lasso_trn_ca, model.matrix(Price~.,data=ca_t)[!trn,-1]) -->

<!-- mse_lasso_ca <- mean((ca_t[!trn,]$Price - pred_lasso_ca)^2) -->
<!-- mse_lasso_ca -->

<!-- # PCR -->

<!-- cv_pcr_ca <- pcr(Price ~ ., data = ca_t[trn,c(-6,-7,-12,-13)], validation = "CV", scale=TRUE, segments=cvsegments(nrow(ca_t[trn,]),10)) -->
<!-- mse_cv <- MSEP(cv_pcr_ca) -->
<!-- M <- mse_cv$comps[which.min(as.data.frame(mse_cv$val)[1,])] -->

<!-- pcr_trn_ca <- pcr(Price ~., data = ca_t[trn,c(-6,-7,-12,-13)], scale=TRUE, M=M) -->
<!-- pred_pcr_ca <- predict(pcr_trn_ca, model.matrix(Price ~ .,data=ca_t[,c(-6,-7,-12,-13)])[!trn,-1]) -->

<!-- mse_pcr_ca <- mean((ca_t[!trn,]$Price - pred_pcr_ca)^2) -->
<!-- mse_pcr_ca -->

<!-- # PLS -->

<!-- cv_pls_ca <- plsr(Price ~ ., data = ca_t[trn, c(-6,-7,-12,-13)], validation = "CV", scale=TRUE,segments=cvsegments(nrow(ca_t[trn,]),10)) -->
<!-- mse_cv <- MSEP(cv_pls_ca) -->
<!-- M <- mse_cv$comps[which.min(as.data.frame(mse_cv$val)[1,])] -->

<!-- pls_trn_ca <- plsr(Price ~., data = ca_t[trn,c(-6,-7,-12,-13)], scale=TRUE, M=M) -->
<!-- pred_pls_ca <- predict(pls_trn_ca, model.matrix(Price ~ .,data=ca_t[,c(-6,-7,-12,-13)])[!trn,-1]) -->

<!-- mse_pls_ca <- mean((ca_t[!trn,]$Price - pred_pls_ca)^2) -->
<!-- mse_pls_ca -->

<!-- ``` -->

<!-- For Colorado I only needed to remove zipcode to solve the collinearity issues. -->

<!-- ```{r} -->

<!-- # Colorado -->

<!-- n <- nrow(co) -->
<!-- folds <- sample(seq_len(k), n, replace = k != n) -->

<!-- # Base Model -->

<!-- base_lm_co <- lm(Price~.,data=co_t) -->
<!-- aliasMatrixCO <- alias(base_lm_co)$Complete -->

<!-- ## COMPUTE ERROR HERE -->

<!-- # Ridge -->

<!-- cv_ridge_co <- cv.glmnet(model.matrix(Price~.,data=co_t)[,-1],unlist(co_t[,1]),alpha=0,lambda=lambda,foldid=folds) -->

<!-- # Lasso -->

<!-- cv_lasso_co <- cv.glmnet(model.matrix(Price~.,data=co_t)[,-1],unlist(co_t[,1]),alpha=1,lambda=lambda,foldid=folds) -->

<!-- # PCR -->



<!-- # PLS -->



<!-- ``` -->

<!-- For LA I removed Zipcode and Bathroom and this solved the perfect collinearity issues. -->

<!-- ```{r} -->

<!-- # Los Angeles -->

<!-- n <- nrow(la) -->
<!-- folds <- sample(seq_len(k), n, replace = k != n) -->

<!-- # Base Model -->

<!-- base_lm_la <- lm(Price~.,data=la_t) -->
<!-- aliasMatrixLA <- alias(base_lm_la)$Complete -->

<!-- ## NEED TO COMPUTE CV ERROR HERE -->



<!-- # Ridge -->

<!-- cv_ridge_la <- cv.glmnet(model.matrix(Price~.,data=la_t)[,-1],unlist(la_t[,1]),alpha=0,lambda=lambda,foldid=folds) -->

<!-- # Lasso -->

<!-- cv_lasso_la <- cv.glmnet(model.matrix(Price~.,data=la_t)[,-1],unlist(la_t[,1]),alpha=1,lambda=lambda,foldid=folds) -->

<!-- # PCR -->



<!-- # PLS -->



<!-- ``` -->

<!-- For Denver I only needed to remove zipcode to solve the collinearity issues. -->

<!-- ```{r} -->

<!-- # Denver -->

<!-- n <- nrow(den) -->
<!-- folds <- sample(seq_len(k), n, replace = k != n) -->

<!-- # Base Model -->

<!-- base_lm_den<- lm(Price~.,data=den_t) -->
<!-- aliasMatrixDEN <- alias(base_lm_den)$Complete -->

<!-- ## NEED TO COMPUTE CV ERROR HERE -->



<!-- # Ridge -->

<!-- cv_ridge_den <- cv.glmnet(model.matrix(Price~.,data=den_t)[,-1],unlist(den_t[,1]),alpha=0,lambda=lambda,foldid=folds) -->

<!-- # Lasso -->

<!-- cv_lasso_den <- cv.glmnet(model.matrix(Price~.,data=den_t)[,-1],unlist(den_t[,1]),alpha=1,lambda=lambda,foldid=folds) -->

<!-- # PCR -->



<!-- # PLS -->



<!-- ``` -->



<!-- # Code for Appendix (if it is needed for appendix in slides and paper) -->

<!-- ```{r show-code, ref.label = all_labels(), echo = TRUE, eval = FALSE} -->

<!-- ``` -->