---
title: "445TestPresentation"
author: 
  - Zach Goldstein
  - Sam Rolsten
  - Andrew Davenport
  - Ryan Lang
date: "`r Sys.Date()`"
output:
  beamer_presentation:
    theme: "PaloAlto"
    colortheme: "spruce"
    fonttheme: "structurebold"
---


```{r setup, include=FALSE}
# These knit options prevent weird warnings from showing up when knitting the presentation
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  cache = FALSE)

# load in libraries here
library(ggplot2)
library(tidyverse)
library(tinytex)
```

# This makes a new slide (one #)

- These dashes make bullet points for the slide
  1. You can also make further indented bulletted points
  2. They can be numbered
  - Or they can be dashed
  
> You can also use ">" to make italicized text 
  
# This makes a new slide (one #)

## You can also make subheadings in the slides that make almost a "table"

- You can use this to add more organized info
- This is helpful to add tables/graphics that you want to title
  
```{r global_eda_setup, include=FALSE}
# Load libraries
library(dplyr)
library(nflfastR)
library(ggplot2)
library(tidyverse)
library(gridExtra)


# Load data
tmp_pbp <- load_pbp()
pbp_all <- load_pbp(1999:2025)


# Build main filtered datasets
fourth <- pbp_all %>%
filter(down == 4, !is.na(fourth_down_converted))


fourth_all <- pbp_all %>%
filter(down == 4) %>%
mutate(
decision = case_when(
play_type == "punt" ~ "Punt",
play_type == "field_goal" ~ "Field Goal",
play_type %in% c("run", "pass", "qb_kneel", "qb_spike", "scramble", "sack") ~ "Go For It",
TRUE ~ NA_character_
)
) %>%
filter(!is.na(decision))


# Helper functions
make_decision_share <- function(data, var, bins = 20) {
  var_sym <- rlang::ensym(var)
  data %>%
    mutate(
      var_bin = if (is.numeric(.data[[rlang::as_string(var_sym)]])) {
        cut(.data[[rlang::as_string(var_sym)]], bins)
        } else {
          as.factor(.data[[rlang::as_string(var_sym)]])
          }) %>%
    group_by(var_bin, decision) %>%
    summarise(n = n(), .groups = "drop") %>%
    group_by(var_bin) %>%
    mutate(share = n/sum(n)) %>%
    ungroup()
  }


plot_decision_share <- function(df, var) {
  ggplot(df, aes(x = var_bin, y = share, fill = decision)) +
    geom_col(position = "fill") +
    labs( title = paste("4th Down Decision Percentages by", var),
          x = var,
          y = "Percent of Decisions",
          fill = "Decision Type") +
    scale_y_continuous(labels = scales::percent) +
    theme_minimal(base_size = 18) + # larger base font
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1, size = 16),
      axis.text.y = element_text(size = 16),
      plot.title = element_text(size = 20))}


# Pre-compute grouped data
by_yl <- make_decision_share(fourth_all, yardline_100, bins = 20)
by_dist <- make_decision_share(fourth_all, ydstogo, bins = 12)
by_score <- make_decision_share(fourth_all, score_differential, bins = 20)
by_q <- make_decision_share(fourth_all, qtr)
```

# Problem Description


## Data Insight
- Wanted to predict what NFL teams should do on 4th down
- Can either punt, kick a field goal, or go for it
- From those options, if they go for it, can either run or pass the ball
- Selected the **nflfastR** data set
- Filtered down to only 4th down plays from 1999-2025
- Split the data into training set (1999-2019) and test set (2021-2025)


# Data Pre-processing


## Overview of Cleaning & Variable Selection
- Loaded full play-by-play dataset (1999–2025) using **nflfastR**.
- Selected key predictors:
- Game context: season, week, quarter, game clock, score differential
- Field context: yardline_100, ydstogo, posteam/defteam timeouts
- Play descriptors: rush, pass, penalty, EPA/WPA, success
- Created a unified **decision** variable:

# Data Pre-processing

## Additional Feature Engineering
- Added intuitive indicators:
  - `close_game` (within 7 points)
  - `short_to_go` (<=  3 yards)
  - `must_go` (trailing, < 5 minutes in 4Q)
  
- Binned variables to reduce model memorization:
  - `yardline_zone`: red_zone_own / mid_field / red_zone_opponent
  - `ydstogo_bin`: short / medium / long / very_long

# Conversion Probability by Distance

```{r, echo=FALSE}
# ensure 'fourth' exists
fourth <- pbp_all %>%
filter(down == 4, !is.na(fourth_down_converted))
```



```{r, echo=FALSE, fig.width=5, fig.height=2}
# jitter plot with larger axis labels and smaller output size
ggplot(fourth, aes(
x = ydstogo,
y = as.numeric(fourth_down_converted),
color = factor(fourth_down_converted, labels = c("Failed", "Converted")),
shape = factor(fourth_down_converted, labels = c("Failed", "Converted"))
)) +
geom_jitter(alpha = 0.6, width = 0.3, height = 0.05, size = 3) +
scale_y_continuous(breaks = c(0,1), labels = c("Failed", "Converted")) +
labs(title = "4th Down Conversions by Yards to Go",
x = "Yards to Go", y = "Outcome",
color = "Conversion", shape = "Conversion") +
theme_minimal()
```

- Conversion probability decreases as distance increases.
- Noticeable separation around 4 yards.



# Decision Percentage: Field Position

```{r, echo=FALSE, fig.width=6, fig.height=3}
ggplot(by_yl, aes(x = var_bin, y = share, fill = decision)) +
  geom_col(position = "fill") +
  labs(
    title = "4th Down Decisions by Field Position",
    x = NULL,  # remove x-axis label
    y = "Percent of Decisions",
    fill = "Decision Type"
  ) +
  scale_y_continuous(labels = scales::percent) +
  theme_minimal(base_size = 16) +
  theme(
    plot.title = element_text(size = 18),   # smaller title
    axis.text.x = element_blank(),          # hide x-axis labels
    axis.text.y = element_text(size = 14)
  )
```

- Punt dominates deep in own territory.
- Go-for-it frequency increases past midfield.



# Decision Percentage: Yards to Go

```{r, echo=FALSE, fig.width=6, fig.height=3}
ggplot(by_dist, aes(x = var_bin, y = share, fill = decision)) +
  geom_col(position = "fill") +
  labs(
    title = "4th Down Decisions by Yards to Go",
    x = NULL,
    y = "Percent of Decisions",
    fill = "Decision Type"
  ) +
  scale_y_continuous(labels = scales::percent) +
  theme_minimal(base_size = 16) +
  theme(
    plot.title = element_text(size = 18),
    axis.text.x = element_blank(),
    axis.text.y = element_text(size = 14)
  )
```

- Longer distances strongly favor punts.
- Short-yardage situations increase aggressiveness.



# Decision Percentage: Score Differential

```{r, echo=FALSE, fig.width=6, fig.height=3}
ggplot(by_score, aes(x = var_bin, y = share, fill = decision)) +
  geom_col(position = "fill") +
  labs(
    title = "4th Down Decisions by Score Differential",
    x = NULL,
    y = "Percent of Decisions",
    fill = "Decision Type"
  ) +
  scale_y_continuous(labels = scales::percent) +
  theme_minimal(base_size = 16) +
  theme(
    plot.title = element_text(size = 18),
    axis.text.x = element_blank(),
    axis.text.y = element_text(size = 14)
  )
```

- Trailing teams are more aggressive.
- Leading teams favor safer options.
